---
title: "Preprocessing"
output: html_document
date: "2024-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Feature Creation

**Load Dataset**

```{r}
df_raw <- readr::read_csv("spain_082019_tweets_csv_hashed.csv")
tail(df_raw)

```

**Choose selection of columns and turn hashtags into a list of vectors**

```{r}
df <- subset(df_raw, select=c(userid, user_reported_location, follower_count, following_count, account_creation_date, account_language, tweet_language, quote_count, reply_count, like_count, retweet_count, hashtags))

# Convert hashtags into a list of vectors
df$hashtags <- lapply(df$hashtags, function(h) {
  h <- strsplit(h, split = "', '")[[1]]  # Split on delimiter
  gsub("\\[|\\]|'", "", h)  # Remove brackets and quotes
})

df$hashtags[df$hashtags == ""] <- NA

df
```

**Calculate the most frequent used hashtag per user**

```{r}
mode <- function(x) {
  if(all(is.na(x))){
    NA
  }else {freq_table <- table(x)
  max_freq <- max(freq_table)
  modes <- names(freq_table[freq_table == max_freq])
  modes[1]}
  
}

user_hashtags <- data.frame(
  userid = rep(df$userid, lengths(df$hashtags)),
  hashtags = unlist(df$hashtags)          
)


user_mode_hashtags <- aggregate(hashtags ~ userid, data = user_hashtags, FUN = mode, na.action =NULL )
user_mode_hashtags
```

```{r}
# Standardize by converting to lowercase
user_mode_hashtags$hashtags <- tolower(user_mode_hashtags$hashtags)

# Manual standardization for specific cases
user_mode_hashtags$hashtags[user_mode_hashtags$hashtags %in% c("sánchezseesconde", "sanchezseesconde")] <- "sanchezseesconde"
user_mode_hashtags$hashtags[user_mode_hashtags$hashtags %in% c("stopokupas", "stopokupas")] <- "stopokupas"
user_mode_hashtags$hashtags[user_mode_hashtags$hashtags %in% c("sánchezmentiroso", "sanchezmentiroso")] <- "sanchezmentiroso"
user_mode_hashtags$hashtags[user_mode_hashtags$hashtags %in% c("stopsoe", "stopsanchez", "sto")] <- "stopsanchez"
```

Most common are: niobreroniespañol - 31 sanchezseesconde - 30 stopokupas - 24 pedroselofunde - 13 decretazosánchez - 12 stopsanchez - 9 nohablamoshacemos - 7 ventealpp - 6 casadoconespaña - 6 objetivocasado - 6 prisiónpermanenterevisable - 5 vuelvezapatero - 5

Make categories about sentiment of hashtags based on my understanding of their meanings:

#Anti-Sánchez/PSOE: Hashtags criticizing or opposing the current government or Pedro Sánchez. sanchezseesconde sanchezmentiroso despídetedesánchez elsanchismoesruina decretazosánchez decretazotegi stopokupas stopsanchez stopsoe

#Pro-PP/Casado: Hashtags supporting the Popular Party (PP) or its leader, Pablo Casado. casadoa3n casadoconespaña casadoconfjl objetivocasado pablocasadoeh ventealpp másppmáscultura másppmásigualdad másppmenosimpuestos valorseguro

#Civic/General Issues: Broader topics not directly tied to political parties but relevant to societal concerns. prisiónpermanenterevisable democraciarealya libertadencataluña quenotelacuelen viveponiente republicaadelante

#Media and Events: Hashtags referencing debates, media appearances, or popular culture. l6ndebatea7 debatertve micasaelecciones lavozasaltos3 lavozsemifinal éraseunarvezcasado juegodeescaños felizjueves

#Other: Hashtags that do not fit neatly into other categories or have unclear context. niobreroniespañol rumbournasarv miprimerTweet túaboadillayoamajadahonda vuelvezapatero nuevovídeodelpsoe

```{r}
user_mode_hashtags$hashtags
```


```{r}
# Define a function to categorize hashtags
categorize_hashtag <- function(hashtag) {
  if (hashtag %in% c("sanchezseesconde", "sanchezmentiroso", "despídetedesánchez",
                     "elsanchismoesruina", "decretazosánchez", "decretazotegi",
                     "stopokupas", "stopsanchez", "stopsoe")) {
    return("Anti-Sanchez/PSOE")
  } else if (hashtag %in% c("casadoa3n", "casadoconespaña", "casadoconfjl",
                            "objetivocasado", "pablocasadoeh", "ventealpp",
                            "másppmáscultura", "másppmásigualdad",
                            "másppmenosimpuestos", "valorseguro")) {
    return("Pro-PP/Casado")
  } else if (hashtag %in% c("prisiónpermanenterevisable", "democraciarealya",
                            "libertadencataluña", "quenotelacuelen",
                            "viveponiente", "republicaadelante")) {
    return("Civic/General Issues")
  } else if (hashtag %in% c("l6ndebatea7", "debatertve", "micasaelecciones",
                            "lavozasaltos3", "lavozsemifinal",
                            "éraseunarvezcasado", "juegodeescaños", "felizjueves")) {
    return("Media and Events")
  } else if (hashtag %in% c("pedroselofunde", "stopsanchez", "stopokupas",
                            "decretazosánchez", "elsanchismoesruina", "sanchezmentiroso", 
                            "ghdúo", "lavozasaltos3", "micasaelecciones", "eldebatelv")) {
    return("Anti-Socialist Campaign")
  } else if (hashtag %in% c("porlaconciliaciónreal", "nohablamoshacemos", "falconviajes",
                            "elppeseconomía", "assassinscreed", "agua")) {
    return("Social and Economic Topics")
  } else if (hashtag %in% c("niobreroniespañol", "rumbournasarv", "miprimerTweet",
                            "túaboadillayoamajadahonda", "vuelvezapatero",
                            "nuevovídeodelpsoe")) {
    return("Cultural/Personal Topics")
  } else if (hashtag %in% c("juegodetronos", "endirecto", "felizjueves", 
                            "concalma", "deseo", "ghdúo9a")) {
    return("Lifestyle and Fun")
  } else {
    return("Uncategorized")
  }
}

# Apply the function to create a new column


user_mode_hashtags$hashtag_category <- sapply(user_mode_hashtags$hashtags, categorize_hashtag)

table(user_mode_hashtags$hashtag_category)

sum(is.na(user_mode_hashtags$hashtag_category))

(user_mode_hashtags)

```

## Create basic user statistics

```{r}
#group by user id and create aggregates for 

mode2 <- function(x) {
  x <- x[!is.na(x)] # Remove NA values
  if (length(x) == 0) return(NA) # Return NA for empty groups
  freq_table <- table(x)
  names(freq_table)[which.max(freq_table)]
}

users <- unique(df$userid)
user_location <- aggregate(user_reported_location ~ userid, data=df, FUN = mode2)
user_follower <- aggregate(follower_count ~ userid, data=df, FUN = mode2)
user_following <- aggregate(following_count ~ userid, data=df, FUN = mode2)
user_acc_creation <- aggregate(account_creation_date ~ userid, data=df, FUN = mode2)
user_acc_language <- aggregate(account_language ~ userid, data=df, FUN = mode2)
user_tweet_language <- aggregate(tweet_language ~ userid, data=df, FUN = mode2)
```

## Create aggregated user statistics

```{r}
aggregated_stats <- aggregate(cbind(quote_count, reply_count, like_count, retweet_count) ~ userid, data=df, FUN = mean)

aggregated_stats
```

```{r}
# Create a named vector for the standardized mapping
location_map <- c(
  "Islas Canarias, España" = "Canary Islands, Spain",
  "Community of Madrid, Spain" = "Madrid, Spain",
  "Burgos, España" = "Burgos, Spain",
  "Madrid, Spain" = "Madrid, Spain",
  "Alicante, Spain" = "Alicante, Spain",
  "El Escorial, España" = "El Escorial, Spain",
  "Seville, Spain" = "Seville, Spain",
  "Madrid" = "Madrid, Spain",
  "Córdoba, España" = "Córdoba, Spain",
  "Rome, Lazio" = "Rome, Italy",
  "España" = "Spain",
  "Cantabria, España" = "Cantabria, Spain",
  "Valencia, Spain" = "Valencia, Spain",
  "Madrid, Comunidad de Madrid" = "Madrid, Spain",
  "Soto del Real, España" = "Soto del Real, Spain",
  "Fuenlabrada, Spain" = "Fuenlabrada, Spain",
  "Barcelona, Spain" = "Barcelona, Spain",
  "Barcelona, España" = "Barcelona, Spain",
  "Segovia, Spain" = "Segovia, Spain",
  "Ribadeo, Spain" = "Ribadeo, Spain",
  "Getafe, España" = "Getafe, Spain",
  "Santander" = "Santander, Spain",
  "Comunidad de Madrid, España" = "Madrid, Spain",
  "Cádiz-Sevilla" = "Cádiz, Spain",
  "Spain" = "Spain",
  "Toledo, España" = "Toledo, Spain",
  "Pontevedra, España" = "Pontevedra, Spain",
  "Granada, Spain" = "Granada, Spain",
  "Zaragoza, España" = "Zaragoza, Spain",
  "Principado de Asturias, España" = "Asturias, Spain",
  "Cáceres, España" = "Cáceres, Spain",
  "Lituania" = "Lithuania",
  "Málaga, España" = "Málaga, Spain",
  "Sevilla" = "Seville, Spain",
  "Sevilla, España" = "Seville, Spain",
  "Venezuela" = "Venezuela",
  "Detrás de ti" = NA,
  "Toledo" = "Toledo, Spain",
  "Cuenca" = "Cuenca, Spain",
  "Malaga, Spain" = "Málaga, Spain",
  "Santander, Spain" = "Santander, Spain",
  "Cadiz, Spain" = "Cádiz, Spain",
  "Madrid - Plymouth" = "Madrid, Spain",
  "Santiago de Compostela, Spain" = "Santiago de Compostela, Spain",
  "Gijón, España" = "Gijón, Spain",
  "Jaen, Spain" = "Jaén, Spain",
  "Málaga" = "Málaga, Spain",
  "Albacete, Spain" = "Albacete, Spain",
  "Santa Cruz de Tenerife-Madrid" = "Santa Cruz de Tenerife, Spain",
  "Móstoles, España" = "Móstoles, Spain",
  "Lleida, España" = "Lleida, Spain",
  "Lisboa, Portugal" = "Lisbon, Portugal",
  "Huesca, España" = "Huesca, Spain",
  "Cádiz, España" = "Cádiz, Spain",
  "Extremadura, España" = "Extremadura, Spain",
  "Ávila, España" = "Ávila, Spain",
  "Santa cruz de tenerife, españa" = "Santa Cruz de Tenerife, Spain",
  "Toledo, Spain" = "Toledo, Spain",
  "Getafe, Spain" = "Getafe, Spain",
  "Valencia, España" = "Valencia, Spain"
)

user_location$user_reported_location[is.na(user_location$user_reported_location)] <- "NA"

# Use the mapping to standardize the locations in the vector
user_location$user_reported_location <- location_map[user_location$user_reported_location]
```

### Left join everything to create final dataset

```{r}

final_nodes <- merge(x = aggregated_stats, y = user_location, by = "userid", all.x = TRUE)
final_nodes <- merge(x = final_nodes, y = user_follower, by = "userid", all.x = TRUE)
final_nodes <- merge(x = final_nodes, y = user_following, by = "userid", all.x = TRUE)
final_nodes <- merge(x = final_nodes, y = user_acc_creation, by = "userid", all.x = TRUE)
final_nodes <- merge(x = final_nodes, y = user_acc_language, by = "userid", all.x = TRUE)
final_nodes <- merge(x = final_nodes, y = user_tweet_language, by = "userid", all.x = TRUE)
final_nodes <- merge(x = final_nodes, y = user_mode_hashtags, by = "userid", all.x = TRUE)
final_nodes
```

# Create network edges dataset

```{r}
reply_network <- na.omit(subset(df_raw, select=c(userid, in_reply_to_userid)))
reply_network <- subset(reply_network, in_reply_to_userid %in% users)

retweet_network <- na.omit(subset(df_raw, select=c(userid, retweet_userid)))
retweet_network <- subset(retweet_network, retweet_userid %in% users)

reply_network
retweet_network

```

## Basic Visualizations
Follower counts:
```{r}
# You can get summary statistics of the follower count
table(final_nodes$follower_count)
final_nodes$follower_count <- as.numeric(final_nodes$follower_count)
final_nodes$following_count <- as.numeric(final_nodes$following_count)

hist(final_nodes$follower_count, 
     main = "Histogram of Follower Counts", 
     xlab = "Follower Count", 
     ylab = "Frequency", 
     breaks = 50,         
     col = "lightblue",     
     border = "white")  

```
A lot of the accounts only have a few followers (0-100): the mean is 67.55

Following counts:
```{r}
# You can get summary statistics of the follower count
final_nodes$following_count <- as.numeric(final_nodes$following_count)
table(final_nodes$following_count)

hist(final_nodes$following_count, 
     main = "Histogram of Following Counts", 
     xlab = "Following Count", 
     ylab = "Frequency", 
     breaks = 50,         
     col = "lightblue",     
     border = "white")  

```

Categories of hashtags:
```{r}
category_counts <- table(final_nodes$hashtag_category)
category_counts
```
```{r}
# Create a bar plot
barplot(category_counts, main = "Distribution of Hashtag Categories", xlab = "Hashtag Categories", ylab = "Frequency", las = 2, col = "lightblue")

```
Anti-Sanchez/PSOE is the most common category.

```{r}
net
```

### Load the edges and join replies and retweets \> engagement dataframe

```{r}
colnames(reply_network) <- c("from", "to")
colnames(retweet_network) <- c("from", "to")

edges <- rbind(reply_network, retweet_network)
```

#Make an igraph of the edges
```{r}
engagement_graph <- snafun::to_igraph(edges, bipartite = FALSE, vertices = final_nodes)
snafun::g_summary(engagement_graph)
```


```{r}
# Remove self-loops in the igraph object
engagement_no_loops <- igraph::simplify(engagement_graph, remove.loops = TRUE, remove.multiple = FALSE)

# Remove isolates
engagement_no_loops_isolates <- snafun::remove_isolates(engagement_no_loops)

# Convert back to a snafun network
net <- snafun::to_network(engagement_no_loops_isolates)
```

```{r}
#Normalize numeric attributes
follower_count <- final_nodes$follower_count
normalized_follower_count <- (follower_count - min(follower_count, na.rm = TRUE)) / 
                             (max(follower_count, na.rm = TRUE) - min(follower_count, na.rm = TRUE))
network::set.vertex.attribute(net, "normalized_follower_count", value = normalized_follower_count)

following_count <- final_nodes$following_count
normalized_following_count <- (following_count - min(following_count, na.rm = TRUE)) / 
                              (max(following_count, na.rm = TRUE) - min(following_count, na.rm = TRUE))
network::set.vertex.attribute(net, "normalized_following_count", value = normalized_following_count)

```


```{r}
# Assign `userid` as the vertex name for matching
network::set.vertex.attribute(net, "userid", value = final_nodes$userid)

# Set other vertex attributes
network::set.vertex.attribute(net, "user_reported_location", value = final_nodes$user_reported_location)
#network::set.vertex.attribute(net, "follower_count", value = final_nodes$follower_count)
#network::set.vertex.attribute(net, "following_count", value = final_nodes$following_count)
network::set.vertex.attribute(net, "account_language", value = final_nodes$account_language)
network::set.vertex.attribute(net, "hashtag_category", value = final_nodes$hashtag_category)
network::set.vertex.attribute(net, "hashtags", value = final_nodes$hashtags)
```

```{r}
node_degrees <- sna::degree(net, cmode=="indegree")
table(node_degrees)

```


```{r}
# Summarize the updated network
snafun::g_summary(net)

#Determine key nodes
key_nodes <- which(node_degrees > 50)
# Default color for all nodes
node_colors <- rep("skyblue", network::network.size(net))  
# Assign "red" to nodes with degree > threshold
node_colors[key_nodes] <- "red"

#Highlight bridge nodes
bridge_nodes <- (sna::betweenness(net)) > 1000
node_colors[bridge_nodes] <- "yellow"

network::plot.network(net, 
                      main = "Network Plot with Highlighted Nodes", 
                      vertex.cex = 1.5,         # Node size
                      vertex.col = node_colors, # Use the color vector
                      edge.col = "gray",        # Edge color
                      mode = "fruchtermanreingold")  # Layout   

# Add a legend
legend("bottomleft",                   # Position of the legend
       legend = c("Key Nodes (Degree > 50)", "Bridge Nodes (Betweenness > 1000)", "Other Nodes"),
       col = c("red", 'yellow', "skyblue"),    # Corresponding colors
       pch = 19,                    # Symbol (circle)
       pt.cex = 1,                 # Size of the symbols
       bty = "n")  
```
#Determine communities with walktrap algorithm
```{r}
graph <- snafun::to_igraph(net)
walktrap <- snafun::extract_comm_walktrap(graph)

plot(walktrap, graph, 
     vertex.label = NA,
     vertex.size = 5,         
     edge.width = 1,         
     edge.color = "grey",
     edge.arrow.size = 0.2)
```

#Add communities as vertex attribute
```{r}
# Extract community membership from the walktrap result
membership <- igraph::membership(walktrap)

table(membership)
# Remove names from the community membership vector
membership <- as.vector(membership)

# Add the community membership as a vertex attribute to the network object
network::set.vertex.attribute(net, "community", value = membership)

# Check if the community attribute is added correctly
print(network::get.vertex.attribute(net, "community"))
```
### Data analysis (Research Rationale)

(about 500 words) – 1 POINTS \* Why is the ERGM/GERM/TERGM suitable for your research questions? \* Why did you pick the specific terms you picked in the model in order to test your hypotheses? \* Are there other methods to address these questions? If yes, why are the methods you chose better for this case?

> Justify the dataset’s relevance to your research question by linking it to the structural terms you’re analyzing.

> ERGM is particularly suited because it allows for modeling of complex structural dependencies, such as reciprocity and transitivity, which are expected in a coordinated network.

> **Choice of Terms**: Justify each term selection in detail:

### Baseline model

```{r}
m0 <- ergm::ergm(net ~ edges)
summary(m0)
```


```{r}
m1 <- ergm::ergm(net ~ edges + nodecov("community"))
summary(m1)
```

```{r}
gof_results <- ergm::gof(m1)
plot(gof_results)
```

```{r}
network::list.vertex.attributes(net)
```




```{r}
m2 <- ergm::ergm(net ~ edges 
                 + nodecov("community") 
                 + nodematch("hashtag_category"))
  
  
summary(m2)
```
```{r}
gof_results <- ergm::gof(m2)
plot(gof_results)
```
```{r}
#look at what value is meaningful for the indegree
max_in_degree <- max(sna::degree(net, cmode = "indegree"))
print(max_in_degree)

in_degree <- sna::degree(net, cmode = "indegree")
table(in_degree)

```

```{r}
m3 <- ergm::ergm(net ~ edges 
                 + nodecov("community") 
                 + nodematch("hashtag_category")
                 + idegree(d=0))
  
  
summary(m3)
```
```{r}
m3 <- ergm::ergm(net ~ edges 
                 + nodecov("community") 
                 + nodematch("hashtag_category")
                 + idegree(d=0),
                 control = ergm::control.ergm(
                   MCMC.samplesize = 15000,
                   MCMC.burnin = 5000,
                   MCMLE.maxit = 20,
                   parallel = 4,
                   parallel.type = "PSOCK"))
  
  
summary(m3)
```

```{r}
# Clear the current plot window
# Adjust margins
par(mar = c(3, 3, 1, 1))  # You can experiment with different values here

```


```{r}
#ergm::mcmc.diagnostics(m3)
gof_results <- ergm::gof(m3)
plot(gof_results)

```


```{r}
m4 <- ergm::ergm(net ~ edges 
                 + nodecov("community") 
                 + nodematch("hashtag_category")
                 + idegree(d=0)
                 + mutual,
                 control = ergm::control.ergm(
                   MCMC.samplesize = 15000,
                   MCMC.burnin = 5000,
                   MCMLE.maxit = 20,
                   parallel = 4,
                   parallel.type = "PSOCK"))
  
  
summary(m4)
```
```{r}
gof_m4 <- ergm::gof(m4)
plot(gof_m4)
```


```{r}
m5 <- ergm::ergm(net ~ edges 
                 + nodecov("community") 
                 + nodematch("hashtag_category")
                 + mutual
                 + gwesp(0.25, fixed = TRUE),
                 control = ergm::control.ergm(
                   MCMC.samplesize = 15000,
                   MCMC.burnin = 5000,
                   MCMLE.maxit = 20,
                   parallel = 4,
                   parallel.type = "PSOCK"))
  
  
summary(m5)
```
```{r}
gof_m5 <- ergm::gof(m5)
plot(gof_m5)
```

```{r}
ergm:: search.ergmTerms("star")
```

```{r}
#Did not run
m5 <- ergm::ergm(net ~ edges 
                 + nodecov("community") 
                 + nodematch("hashtag_category")
                 + mutual
                 + gwesp(0.25, fixed = TRUE)
                 + istar(20),
                 control = ergm::control.ergm(
                   MCMC.samplesize = 15000,
                   MCMC.burnin = 5000,
                   MCMLE.maxit = 20,
                   parallel = 4,
                   parallel.type = "PSOCK"))
  
  
summary(m5)
```
#More endogenous terms
```{r}
m6 <- ergm::ergm(net ~ edges 
                 + nodecov("community") 
                 + nodematch("hashtag_category")
                 + nodeicov("follower_count")
                 + idegree(d=0)
                 + mutual,
                 control = ergm::control.ergm(
                   MCMC.samplesize = 15000,
                   MCMC.burnin = 5000,
                   MCMLE.maxit = 20,
                   parallel = 4,
                   parallel.type = "PSOCK"))
  
  
summary(m6)
```
```{r}
gof_m6 <- ergm::gof(m6)
plot(gof_m6)
```

```{r}

m6 <- ergm::ergm(net ~ edges 
                 + nodecov("community") 
                 + nodematch("hashtag_category")
                 + nodeicov("follower_count")
                 + odegree(0)
                 + mutual
                 + gwesp(0.25, fixed = TRUE),
                 control = ergm::control.ergm(
                   MCMC.samplesize = 15000,
                   MCMC.burnin = 5000,
                   MCMLE.maxit = 20,
                   parallel = 4,
                   parallel.type = "PSOCK"))
  
  
summary(m6)
```

```{r}

m7 <- ergm::ergm(net ~ edges 
                 + nodematch("community") 
                 + nodeicov("follower_count")
                 + idegrange(from = 50, to = +Inf)
                 + mutual
                 + gwdsp(2, fixed = TRUE),
                 control = ergm::control.ergm(
                   MCMC.samplesize = 15000,
                   MCMC.burnin = 5000,
                   MCMLE.maxit = 20,
                   parallel = 4,
                   parallel.type = "PSOCK"))
  
  
summary(m7)
```

```{r}
gof_m7 <- ergm::gof(m7)
plot(gof_m7)
```

```{r}
betw_f <- function(x, directed = FALSE) {  # directed change to true/false
  x <- snafun::fix_cug_input(x, directed = directed) #always have this line
  
}

cug_flomar_betw <- sna::cug.test(flomarriage, mode = "graph", #graph=undirected
                                 FUN = betw_f, 
                                 cmode = "edges", reps = 1000)
cug_flomar_betw
sna::plot.cug.test(cug_flomar_betw)
```



```{r}
m8 <- ergm::ergm(net ~ edges 
                 + nodematch("community") 
                 + nodeicov("follower_count")
                 + idegrange(from = 18, to = +Inf)
                 + mutual
                 + gwdsp(2, fixed = TRUE),
                 control = ergm::control.ergm(
                   MCMC.samplesize = 15000,
                   MCMC.burnin = 5000,
                   MCMLE.maxit = 20,
                   parallel = 4,
                   parallel.type = "PSOCK"))
  
  
summary(m8)
```

```{r}
gof_m8 <- ergm::gof(m8)
plot(gof_m8)
```

```{r}
ergm::mcmc.diagnostics(m8)
```


```{r}
m9 <- ergm::ergm(net ~ edges 
                 + nodematch("community") 
                 + nodematch("hashtag_category")
                 + idegrange(from = 50, to = +Inf)
                 + mutual
                 + gwdsp(2, fixed = TRUE),
                 control = ergm::control.ergm(
                   MCMC.samplesize = 15000,
                   MCMC.burnin = 5000,
                   MCMLE.maxit = 20,
                   parallel = 4,
                   parallel.type = "PSOCK"))
  
  
summary(m9)
```

